Making Sequential Programs Deterministic

### 2019-04-06 experiments.tex
expts edit

We modified Randoop by deleting tests that were skipped by its buildfile,
because the NonDex Gradle plugin does not respect those settings.
After that, NonDex ran without problems on Randoop.
NonDex reported no flaky tests.
%\todo{Revise the following text.
%  I renumbered the bugs without realizing that this section refers to them
%  by number, and I don't know which one these refer to any
%  more.
%  Please use cross-references or macros.  Also, this needs to be expanded
%  to account for the additional bugs.}
We investigated the tests covering code that \theDeterminismChecker reported as nondeterministic.
\bugHashSet and \bugClasspath have corresponding test cases whereas \bugTimestampOutput and 
\bugTimestampOutput do not.
As NonDex is a \textit{flaky test} detection tool, detecting \bugTimestampOutput and \bugTimestampOutput 
is outside of its scope.
The reason for nondeterminism in \bugClasspath was a call to the \<System.getProperty()> method which
is not modeled by NonDex.

@@

    \begin{subfigure}[b]{0.95\textwidth}
        \begin{verbatim}
    protected SimpleDataEvent createNextEvent() {
        for (Entry<String, FieldType> entry : fields.entrySet()) {
            ...
        }
        ...
    }
        \end{verbatim}
        \caption{Annotated source code from ActionGenerator.}
        \label{code-actiongenerator}
    \end{subfigure}    
    \caption{Errors detected by \TheDeterminismChecker.}
    \label{fig:nondex-source}
\end{figure}

@@

In reflectasm (\cref{code-reflectasm}), we changed the type of field
\<declaredFields> from \<Field[]> to \<@Det Field @Det []>.  That type
means a deterministic array of deterministic \<Field>, analogously to
\codeid{@Det List<@Det Field>}.  Then, \theDeterminismChecker issued a
warning at the assignment, because \<getDeclaredMethods> returns \<@Det
Field @Det []>, which is an order-nondeterministic array of deterministic
\<Field>s.

@@

Four of the seven flaky tests (two in \<achilles>, one each in \<jackrabbit-oak> and \<togglz>)
were caused by a race condition, which \theDeterminismChecker
cannot detect.  (This is a strength of DeFlaker over \theDeterminismChecker.)
We were unable to build \<togglz> and \<nutz> which had one flakiness issue each.
However, we extracted the source code causing the flakiness in these repositories into test cases
after looking at the corresponding issues on Github. The bug in \<togglz> was caused by a copy method that iterated over
an \<OrderNonDet Set> and expected it to be deterministic. \TheDeterminismChecker correctly flagged
an error in the loop that iterated over the \<Set>.
The flakiness in \<nutz> was a result of printing a response received over http.
Since network operations are nondeterministic, we annotated the method in \<nutz> that
returns this response as \<NonDet> which lead \TheDeterminismChecker to report an error at the print statement. 
\TheDeterminismChecker was able to find the source of flakiness in \<checkstyle>. This bug was in a test
case that treated an array returned by \<Class.getDeclaredConstructors()>
as deterministic.

>>>>

We modified Randoop by deleting tests that were skipped by its buildfile,
because the NonDex Gradle plugin does not respect those settings.
After that, NonDex ran without problems on Randoop.
NonDex reported no flaky tests.
%\todo{Revise the following text.
%  I renumbered the bugs without realizing that this section refers to them
%  by number, and I don't know which one these refer to any
%  more.
%  Please use cross-references or macros.  Also, this needs to be expanded
%  to account for the additional bugs.}
We investigated the tests covering code that \theDeterminismChecker reported as nondeterministic.
\bugHashSet and \bugClasspath have corresponding test cases whereas the remainder of the bugs do not.
As NonDex is a \textit{flaky test} detection tool, detecting \bugTimestampOutput and \bugTimestampOutput 
is outside of its scope.
The reason for nondeterminism in \bugClasspath was a call to the \<System.getProperty()> method which
is not modeled by NonDex.

@@

    \begin{subfigure}[b]{0.95\textwidth}
        \begin{verbatim}
    protected SimpleDataEvent createNextEvent() {
        for (Entry<String, FieldType> entry : fields.entrySet()) {
            ...
        }
        ...
    }
        \end{verbatim}
        \caption{Annotated source code from ActionGenerator.}
        \label{code-actiongenerator}
    \end{subfigure}    
    \caption{Errors detected by \TheDeterminismChecker in NonDex benchmarks.}
    \label{fig:nondex-source}
\end{figure}

@@

In reflectasm (\cref{code-reflectasm}), we changed the type of field
\<declaredFields> from \<Field[]> to \<@Det Field @Det []>.  That type
means a deterministic array of deterministic \<Field>, analogously to
\codeid{@Det List<@Det Field>}.  Then, \theDeterminismChecker issued a
warning at the assignment, because \<getDeclaredMethods> returns \<@Det
Field @OrderNonDet []>, which is an order-nondeterministic array of deterministic
\<Field>s.

@@

Four of the seven flaky tests (two in \<achilles>, one each in \<jackrabbit-oak> and \<togglz>)
were caused by a race condition, which \theDeterminismChecker
cannot detect.  (This is a strength of DeFlaker over \theDeterminismChecker.)
We were unable to build \<togglz> and \<nutz> which had one flakiness issue each.
However, we extracted the source code causing the flakiness in these repositories into test cases
after looking at the corresponding issues on Github. The bug in \<togglz> was caused by a copy method that iterated over
an \<OrderNonDet Set> and expected it to be deterministic. \TheDeterminismChecker correctly flagged
an error in the loop that iterated over the \<Set>.
The flakiness in \<nutz> was a result of printing a response received over http.
Since network operations are nondeterministic, we annotated the method in \<nutz> that
returns this response as \<NonDet> which lead \TheDeterminismChecker to report an error at the print statement. 
\TheDeterminismChecker was able to find the source of flakiness in \<checkstyle>. This bug was in a test
case that treated an array returned by \<Class.getDeclaredConstructors()>
as deterministic. This is erroneous because \<getDeclaredConstructors> returns an order nondeterministic array.
